Planetary perturbers: Flaring star-planet interactions in Kepler and TESS

### 2023-10-05 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger flares on the stellar surface. These flares are not expected to appear different from intrinsic stellar flares individually. However, as we expect planet-induced flares to be associated with the planet's orbital period $P_{\rm orb}$, we can identify their presence from a clustering of flares in phase with the planet's orbital phase. We used the excellent phase coverage obtained from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase.

@@

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. The past two decades revealed that planets can occur in orbits of only a few stellar radii~\citep[e.g., ][]{sanchis-ojeda2014study} that they could almost be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions (SPI) that emphasizes that they are bidirectional, affecting both planet, and star. Specifically, the effects on the star pose challenges, but also opportunities for our understanding of star-planet systems:

@@

\subsection{Searching for global changes in activity}
\label{sec:intro:global}
Measuring magnetic star planet-interaction is notoriously difficult in individual systems because a non-detection may not mean an absence of the interaction, but merely a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive, poppenhaeger2010coronal}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces selection bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability~\citep{poppenhaeger2011correlation}, but that does not imply that SPI quenches activity. Further, tidal interactions can induce global increase in activity~\citep{ilic2022tidal}, which adds a physical ambiguity to the interpretation~(see also Section~\ref{sec:discussion:tidal}).

@@

\section*{Acknowledgements}

This paper includes data collected by the TESS mission, which are publicly available from the Mikulski Archive for Space Telescopes (MAST).
Funding for the TESS mission is provided by NASA’s Science Mission directorate. 
This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions participating in the {\it Gaia} Multilateral Agreement.
KP, JC and NI acknowledge funding from the German \textit{Leibniz Community} under grant P67/2018.
%follow instructions here https://gea.esac.esa.int/archive/documentation/GEDR3/Miscellaneous/sec_credit_and_citation_instructions/
\section*{Data Availability}
All data used in this study are publicly available in their respective archives.
This study used the reproducibility software \href{https://github.com/showyourwork/showyourwork}{showyourwork}
\citep{luger2021mappinga} to create all figures, format all tables, and compile the manuscript. Each figure links to the dataset stored on Zenodo (link to be added after acceptance), and to the script that produced the figure in the git repository~\href{https://github.com/ekaterinailin/flaring-spi-paper}{github.com/ekaterinailin/flaring-spi-paper}. The git repository that produces the data stored on Zenodo is publicly available at \href{https://github.com/ekaterinailin/flaring-spi}{github.com/ekaterinailin/flaring-spi}. 
\bibliography{bib}

>>>>

% Abstract with filler text
\begin{abstract}
    In many star-planet systems discovered so far, the innermost planet orbits within only a few stellar radii. In these systems, planets \textcolor{red}{could} become in-situ probes of the extended stellar magnetic field. Because they disturb the field as they move, they are expected to trigger flares on the stellar surface. These flares are not expected to appear different from intrinsic stellar flares individually. However, as we expect planet-induced flares to be associated with the planet's orbital period $P_{\rm orb}$, we can identify their presence from a clustering of flares in phase with the planet's orbital phase. We used the excellent phase coverage obtained from Kepler and the Transiting Exoplanet Survey Satellite to find flaring star-planet systems, compile a catalog of all their flares, and measure how much they cluster in orbital phase.

@@

% Main body with filler text
\section{Introduction}
\label{sec:intro}
One of the most surprising results in exoplanet research is the diversity of star-planet system architectures. The past two decades revealed that planets can occur in orbits of only a few stellar radii~\citep[e.g., ][]{sanchis-ojeda2014study} that they could almost be considered to live inside the stellar atmosphere. At this proximity, the planet exerts force, tidal and magnetic, both short- and long-term, on its host -- a phenomenon largely unknown to the Solar System. These forces are typically subsumed under the term star-planet interactions (SPI) that emphasizes that they are bidirectional, affecting both planet, and star. Specifically, the effects on the star pose challenges, but also opportunities for our understanding of star-planet systems.

@@

\subsection{Searching for global changes in activity}
\label{sec:intro:global}
\textcolor{red}{Even if the instrument's sensitivity is sufficiently high, measuring magnetic star planet-interaction remains notoriously difficult. In individual systems,} a non-detection may not mean an absence of the interaction, but merely a temporary cessation~\citep{shkolnik2005hot, shkolnik2008nature}. Statistical studies searching for changes in global chromospheric and coronal activity indicators in star-planet systems with close-in planets relative to planet-less stars are inconclusive~\citep{kashyap2008extrasolar,scharf2010possible, shkolnik2013ultraviolet, france2018farultraviolet, viswanath2020statistical, krejcova2012evidence, miller2015comprehensive, poppenhaeger2010coronal}. In these studies, it is challenging both to define a control sample to measure the activity against, and to quantify how the way we detect planets introduces selection bias in the activity measure. It is, for instance, easier to detect a planet around a magnetically inactive star, because both radial velocity and transit detections will be less affected by stellar variability~\citep{poppenhaeger2011correlation}, but that does not imply that SPI quenches activity. Further, tidal interactions can induce \textcolor{red}{a} global increase in activity~\citep{ilic2022tidal}, which adds a physical ambiguity to the interpretation~(see also Section~\ref{sec:discussion:tidal}).

@@

\section*{Acknowledgements}
\textcolor{red}{The authors thank the anonymous referee for their thorough criticism of the manuscript, and numerous valuable suggestions. EI would like to thank Antonino Lanza for helpful advice.}
This paper includes data collected by the TESS mission, which are publicly available from the Mikulski Archive for Space Telescopes (MAST).
Funding for the TESS mission is provided by NASA’s Science Mission directorate. 
This work has made use of data from the European Space Agency (ESA) mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by the {\it Gaia} Data Processing and Analysis Consortium (DPAC, \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding for the DPAC
has been provided by national institutions, in particular the institutions participating in the {\it Gaia} Multilateral Agreement.
KP, JC and NI acknowledge funding from the German \textit{Leibniz Community} under grant P67/2018.
%follow instructions here https://gea.esac.esa.int/archive/documentation/GEDR3/Miscellaneous/sec_credit_and_citation_instructions/
\section*{Data Availability}
All data used in this study are publicly available in their respective archives.
This study used the reproducibility software \href{https://github.com/showyourwork/showyourwork}{showyourwork}
\citep{luger2021mappinga} to create all figures, format all tables, and compile the manuscript. Each figure links to the dataset stored on Zenodo (link to be added after acceptance), and to the script that produced the figure in the git repository~\href{https://github.com/ekaterinailin/flaring-spi-paper}{github.com/ekaterinailin/flaring-spi-paper}. The git repository that produces the data stored on Zenodo is publicly available at \href{https://github.com/ekaterinailin/flaring-spi}{github.com/ekaterinailin/flaring-spi}. 
\bibliography{bib}
